Real-Time Activity Classification by Match Filtering using Body-Worn Accelerometers

### 2016-06-20 notes/paper/paper.tex
Intermediate changes
Added more to the paper. Made python script to create timing plot.

\documentclass[journal]{IEEEtran}
\usepackage{blindtext}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{caption}
\usepackage[english]{babel}
\usepackage{pinyin}
\usepackage{natbib}
\usepackage{notoccite}
\usepackage{amssymb}
\usepackage{textcomp}
\usepackage{floatrow}
\usepackage{amsmath}
%
\begin{document}
%
% paper title
\title{Real-Time Activity Classification by Match Filtering using Body-Worn Accelerometers}
%
\author{Craig~Euler, C.T.~Lin, Bryan~Juarez}
%
% make the title area
\maketitle
%
\begin{abstract}
Body worn sensors have been used in the market for many years now.
Many of which are used for activity monitoring or fitness.
In this paper, we make use of a matched-filtering algorithm for real-time activity classification from data obtained with three-axis accelerometers worn by the user.
We achieve invariance to sensor orientation with dimensionality reduction.
We make use of an instance-based learning algorithm to train the device to learn the individual's motion patterns and store that information as activity filters for our matched filter.
We improve computational efficiency with dimensionality reduction, temporal decomposition, and pre-processing of our filters for later use.
An assessment of our computational costs are provided as processing flop counts for the real-time specific parts of the algorithms.
\end{abstract}
%
\section{Introduction}
Various methods for activity classification using body-worn sensors exist.
Custom decision tree, automatically generated decision tree, and artificial neural networks \cite{parkka_ermes_korpipaa_mantyjarvi_peltola_korhonen_2006} have been explored as well as other frequency based classification methods \cite{sharma_purwar_lee_lee_chung_2008}.
In this paper, we make use of a matched-filter method to identify an individual\textquotesingle s activity in real-time.
The matched-filter method with use of thresholding and comparing the correlations of various filters is a simple method to use and have been studied in the past \cite{giannakis_tsatsanis_1990}.
Many of the algorithms and software structure were tailored to computational efficiency to accomidate the resource limitations on the MSP 432 micro-processor.
Having an a priori knowledge of the activity the individual is performing is useful for many current fitness algorithms.
Devices such as the Fit-Bit, Apple watch, Moov, Jawbone, etc. usually have the individual actively input the activity they are performing.

@@

If we let $\textbf{a} = \{a_i\}_{i=0}^{N}$ represent our demeaned acceleration data over a specified time window in three dimensions then the covariance matrix of our acceleration data is represented by the real-symmetric $3 \times 3$ matrix $Q = \textbf{a} \textbf{a}^T$.
The resulting eigen vectors of this matrix are orthogonal and point in the directions where the data varies most [cite from some textbook here].
By projecting our data on the plain defined by the two most dominant eigen vectors, we effectively reduce our representation of our data to two dimensions.

To compute the matched-filter output for two dimensions, we generalize \eqref{matched_filter_eq} by augmenting our reference and data with both dimensions appropriately for normalization and adding together the cross correlations as:
%
\begin{equation} \label{cross_correlation_eq_2}
\widehat{(\textbf{x}^{1,2} \star \textbf{y}^{1,2})}_k^2 = \frac{(\textbf{x}^1 \star \textbf{y}^1)_k + (\textbf{x}^2 \star \textbf{y}^2)_k}{||[ \textbf{x}^1, \textbf{x}^2 ]|| \ || [ \widetilde{\textbf{y}}_k^1, \widetilde{\textbf{y}}_k^2 ] || },
\end{equation}
%
for $ k = 0,1,...,N-M-1 $.\\
%
\begin{equation} \label{matched_filter_eq_2}
MF_{\textbf{x}}^2(\textbf{y}) := \underset{k \in [0, M-N)}{max} \left \{\widehat{(\textbf{x} \star \textbf{y})}^2_k \right \}
\end{equation}
%
This works because both dimensions are always in phase with each other for any given activity which means that the match will happen in sinc.
%
\subsection{Decimation}
Decimation is the first step applied to our data before any additional processing takes place.
Decimation is a process of reducing the sampling rate of a signal.
The process consists of filtering to mitigate any possible aliasing distortion effects followed by a downsampling.
There is a tradeoff between performance and throughput when we decimate our signal, therefore, we use decimation solely for the purpose of improving our throughput.

>>>>

\documentclass[journal]{IEEEtran}
\usepackage{blindtext}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{caption}
\usepackage[english]{babel}
\usepackage{pinyin}
\usepackage{natbib}
\usepackage{notoccite}
\usepackage{amssymb}
\usepackage{textcomp}
\usepackage{floatrow}
\usepackage{amsmath}
%
\begin{document}
%
% paper title
\title{Real-Time Activity Classification by Match Filtering using Body-Worn Accelerometers}
%
\author{Craig~Euler, C.T.~Lin, Bryan~Juarez, Melissa~Flores}
%
\maketitle
%
\begin{abstract}
Body worn sensors have been used in the market for many years now.
Many of which are used for activity monitoring or fitness.
In this paper, we make use of a matched-filtering algorithm for real-time activity classification from data obtained with three-axis accelerometers worn by the user.
We achieve invariance to sensor orientation with dimensionality reduction.
We make use of an instance-based learning algorithm to train the device to learn the individual's motion patterns and store that information as activity filters for our matched filter.
We improve computational efficiency with dimensionality reduction, temporal decomposition, and pre-processing of our filters for later use.
An assessment of our computational costs are provided with runtime experiments on the MSP432 micro-controller.
\end{abstract}
%
\section{Introduction}
Various methods for activity classification using body-worn sensors exist.
Custom decision tree, automatically generated decision tree, and artificial neural networks \cite{parkka_ermes_korpipaa_mantyjarvi_peltola_korhonen_2006} have been explored as well as other frequency based classification methods \cite{sharma_purwar_lee_lee_chung_2008}.
In this paper, we make use of a matched-filter method to identify an individual\textquotesingle s activity in real-time.
The matched-filter method with use of thresholding and comparing the correlations of various filters is a simple method to use and have been studied in the past \cite{giannakis_tsatsanis_1990}.
Many of the algorithms and software structure were tailored to computational efficiency to accomidate the resource limitations on the MSP 432 micro-processor.
Having an a priori knowledge of the activity the individual is performing is useful for many current fitness algorithms.
Devices such as the Fit-Bit, Apple watch, Moov, Jawbone, etc. usually have the individual actively input the activity they are performing.

@@

If we let $\textbf{a} = \{a_i\}_{i=0}^{N}$ represent our demeaned acceleration data over a specified time window in three dimensions then the covariance matrix of our acceleration data is represented by the real-symmetric $3 \times 3$ matrix $Q = \textbf{a} \textbf{a}^T$.
The resulting eigen vectors $V = (\textbf{v}^1,\textbf{v}^2,\textbf{v}^3)$ for the respective eigen values $\lambda^1 \geq \lambda^2 \geq \lambda^3$ of this matrix are orthogonal and point in the directions where the data varies most [cite from some textbook here].
%
\begin{equation} \label{project_eq}
   \textbf{a}' = V^T \textbf{a}
\end{equation}
%
We now represent our acceleration data as $\textbf{a}'$ after rotating our axis by \eqref{project_eq}. We make the assumption that our data is most dominate in two dimensions i.e. $\lambda^3 << \lambda^2$ so it is safe to assume that
%
\begin{equation} \label{magnitude_eq}
||(a^1,a^2,a^3)|| = ||(a'^1,a'^2,a'^3)|| \sim = (a'^1)^2 + (a'^2)^2
\end{equation}
$ \forall a \in \textbf{a}. $ \\
%
The margin of error in the approximation is not important since we are only interested in processing in two dimensions. Our goal is only to minimized the margin of error so that we are not, by chance, using an axis direction where the data varies least.
By projecting our data onto the plain defined by the two most dominant eigen vectors, we effectively reduce our representation of our data to two dimensions.

To compute the matched-filter output for two dimensions, we generalize \eqref{matched_filter_eq} by augmenting our reference and data with both dimensions appropriately for normalization and adding together the cross correlations as:
%
\begin{equation} \label{cross_correlation_eq_2}
\widehat{(\textbf{x}^{1,2} \star \textbf{y}^{1,2})}_k = \frac{(\textbf{x}^1 \star \textbf{y}^1)_k + (\textbf{x}^2 \star \textbf{y}^2)_k}{||[ \textbf{x}^1, \textbf{x}^2 ]|| \ || [ \widetilde{\textbf{y}}_k^1, \widetilde{\textbf{y}}_k^2 ] || },
\end{equation}
%
for $ k = 0,1,...,N-M-1 $.\\
%
\begin{equation} \label{matched_filter_eq_2}
MF_{\textbf{x}}^2(\textbf{y}) := \underset{k \in [0, M-N)}{max} \left \{\widehat{(\textbf{x}^{1,2} \star \textbf{y}^{1,2})}_k \right \}
\end{equation}
%
This works because both dimensions are always in phase with each other for any given activity and therefore the match will happen in sync.
%
\subsection{Decimation}
Decimation is the first step applied to our data before any additional processing takes place.
Decimation is a process of reducing the sampling rate of a signal.
The process consists of filtering to mitigate any possible aliasing distortion effects followed by a downsampling.
There is a tradeoff between performance and throughput when we decimate our signal, therefore, we use decimation solely for the purpose of reducing our processing cost.
